The K2 Bright Star Survey

### 2019-05-09 paper/ms.tex
misc updates

%% Mark off the abstract in the ``abstract'' environment. 
\begin{abstract}
While the \kepler\ mission was designed to look at tens of thousands of faint stars ($V \gtrsim 12$), brighter stars which saturate the detector are nevertheless some of the most interesting because of the ease with which they can be observed by other instruments and the wealth of knowledge about them that is already available. By considering the unsaturated scattered light `halo' around these stars we retrieve precise light curves of most of the brightest stars in \ktwo fields from Campaign~4 onwards. The halo method is highly agnostic about the cause and form of systematics and we show that it nevertheless it is effective at extracting light curves from both normal and saturated stars. This bright star survey reveals stellar variability ubiquitously, including effects of stellar pulsation, rotation, and binarity. Here we describe our pipeline, and present a catalogue of the bright stars studied, with classifications and parametrizations of their variability and remarks on interesting objects. These light curves are publicly available as a High Level Science Product from the Mikulski Archive for Space Telescopes (MAST). \href{https://github.com/benjaminpope/k2halo}{\color{linkcolor}\faGithub} % git link shamelessly borrowed from Luger!
\end{abstract}

@@

% briefly intro the halo method and halo literature
In order to recover precise photometry of the brightest stars in \ktwo, we have developed two main approaches, `smear' and `halo' photometry. Smear photometry \citep{Pope2016} uses collateral calibration data to recover light curves of stars which were not necessarily conventionally targeted. The more precise method of halo photometry, which is the subject of this paper, uses the broad `halo' of scattered light around a saturated star to recover relative photometry, by constructing a light curve as a linear combination of individual pixel time series and minimizing a Total Variation objective function (TV-min). It has been employed for example on the Pleiades \citep{White2017} and the brightest-ever star on \kepler silicon, Aldebaran \citep{Farr2018}, recovering photometry with a precision nearly that of normally-obtained \ktwo observations of unsaturated stars. Unlike smear, this requires downloading data out to a 12--20~pixel radius around each star, and has accordingly only been possible for stars that were specifically targeted with apertures optimized for this method and for a small number of other stars for which this is fortuitously the case.

% describe contents of paper: code, catalog, categorization, conclusions
In this Paper we describe the complete catalog of long-cadence \ktwo halo light curves which we have made publicly available. We have employed halo photometry on all stars targeted with appropriate apertures, and have done a preliminary characterization of interesting astrophysical variability. We also document the main changes in the halo data reduction pipeline, \texttt{halophot}, with respect to previous releases. These include oscillating red giants, pulsating and quiet main sequence stars, and eclipsing binaries, many of which are among the brightest objects of their type to have been observed with space photometry. We hope that this diverse catalog of light curves will be useful for an equally-diverse range of astrophysical investigations.

@@

% describe saturation
The \kepler detector saturates at a magnitude of $K_p \sim 11.3$ in both long- (30\,min) and short (1\,min)-cadence data, as these both represent sums of 6~s exposures \citep{Gilliland2010}. Beyond this point, additional photons cause excess electrons to `bleed' into adjacent pixels in both directions along the column containing the bright star. This process occurs conservatively and simple aperture photometry (SAP) adding all the flux contained in a window around the bleed column has recovered light curves with precisions close to the photon noise limit of stars as bright as 16~Cyg~AB, $\theta$~Cyg and RR~Lyr \citep[e.g.][]{Kolenberg2011,2013MNRAS.433.1262W,Guzik2016}. In the nominal \kepler mission this was not attempted for all such bright stars, and in \ktwo, the several-pixel spacecraft motion significantly increased the size of the required apertures for SAP photometry of very saturated stars, while also making their systematics more difficult to deal with. While the second-version pixel-level-decorrelation (PLD) pipeline EVEREST~2.0 was able to correct systematics in saturated SAP photometry \citep{everest2}, this is rendered impossible for the very brightest stars whose bleed columns may run to the edge of the detector. Furthermore, bandwidth constraints meant that pixel data were not downloaded for many bright targets in \ktwo. It was therefore necessary to develop a method to recover good light curves from a smaller number of unsaturated pixels in a compact aperture around bright stars: halo photometry. 
% smear
This method differs from smear photometry \citep{Pope2016}, which uses collateral `smear' data to obtain a 1-D spatial profile with $\sim 1/1000$ of the flux on each CCD. While this has also been used for photometry of very bright stars, the `halo' method has the advantage of averaging over many more realizations of the K2 pointing systematics and permitting more significant calibration, achieving a generally higher photometric prevision.

@@

\begin{equation}
    \Lambda_{k,l} \equiv {\sum_i{|f_i - f_{i-l}|^k}},
\end{equation}

@@

\begin{align}
\forall_j w_j &> 0\\
\sum_{i=1}^{N} f_i  &= N
\end{align}

This is a classic convex optimization program with constraints, which we solve with the \texttt{scipy} \citep{scipy} L-BFGS-B nonlinear optimization code \citep{lbfgsb}. $\Lambda_{k,l}$ has analytic derivatives with respect to $w_j$ \citep[calculated with \texttt{autograd}][]{autograd}, and it is therefore extremely fast to optimize and converges well on a global solution. In practice, for computational reasons we optimize over parameters $\tilde{w}_j$ such that $w_j = \text{softmax}(\tilde{w}_j$, as this satisfies the constraint that $\forall_j w_j > 0$, and while this also constrains their sum to be unity, we renormalize $f$ to satisfy its normalization constraint before calculating the objective function and this additional constraint is removed again.

The objective function $\Lambda_{k,l}$ is the $L_k$ norm on a `lagged' finite difference with a lag parameter $l$. For $k = 1$ and $l = 1$, $\Lambda_{1,1}$ is the standard Total Variation objective (TV) used in previous halo papers, and can be seen as the L1 norm on the derivative of $f$ or as a discrete approximation to its arc length. The L2 Variation (L2V) with $k=2$ is sometimes referred to in image processing literature as the `smoothness' regularizer, as it seeks to penalize large gradients without necessarily making them sparse. The lag parameter $l$ allows for flexibility in modelling systematics occurring at different timescales from cadence-to-cadence, and we investigate its effects below. The order parameter $k$ allows for flexibility in how sensitive we are to normally-distributed versus long-tailed noise.

@@

In order to choose the values for $k$ and $l$ in our objective function, we have selected the quiet star 36~Ophiuchi, the lowest-mass main sequence star in the halo sample and one with very little high frequency variability detected or predicted, which was also observed at short cadence. We choose the 6.5~hour Combined Differential Photometric Precision \citep[CDPP,][]{cdpp} as implemented in \texttt{lightkurve} \citep{lightkurve} as a proxy for the `noise' in a lightcurve, with lower being better. We calculate halo lightcurves and their CDPPs for $k \in \{1,2,3\}$, and $l \in [1,50]$ for long cadence and (sparsely) $l \in [1,2500]$ for short cadence data. The results are displayed in Figure~\ref{cdpps}. We find that for long cadence data, the L1 norm (TV) and a lag of 10 provide the best CDPP, which is unsurprising given that this is a slightly shorter lag than the 12~cadence thruster firing period. In this context we can understand the optimum as TV-min suppressing systematics on the same timescale as they occur. On the other hand, for short cadence data, performance at short lags is very poor but the method performs similarly for $k \in \{1,2\}$ and improves very slowly with $l$ for longer lags, and performs very poorly for $k=3$ at all lags.

\begin{figure*}
\plottwo{lag_lc_cdpp.pdf}{lag_sc.pdf}
\caption{Behaviour of long cadence (left) and short cadence (right) 6.5~hour CDPP for the quiet dwarf 36~Ophiuchi as a function of lag parameter $l$. For long cadence this shows a minimum for L1 norm and $l=10$, i.e. for objective function $\Lambda_{1,10}$. For short cadence CDPP continuously improves for higher lags and shows no strong differences between L1 and L2~norms, while L3 performs poorly.}
\label{cdpps}
\end{figure*}

%  improvements
Unlike other methods for calibrating \kepler systematics, other than the value of $l$, no knowledge of the spacecraft motion or the behaviour of an ensemble of other stars is used to inform our algorithm. The method is a self-calibration that is independent of the details of the systematics it is calibrating, operating on the assumption that a single signal is present across many individual time series which otherwise are contaminated by noise. It is therefore likely that significant improvements can be made to the method by including cotrending basis vectors with mean zero and whose weights are allowed to be negative, which would represent systematics which are common to all pixels in the halo aperture and therefore masquerade as signal. Any linear combination of convex objective functions is itself convex, and future extensions to the method could enforce combinations of different lags and orders to better represent systematics occurring on different timescales (e.g. thruster firings, red noise) and with different levels of smoothness.

@@

% do k2 smear after that
The sample of K2 bright stars presented here only includes those with halo apertures, but while some others are available conventionally, many were not assigned target pixels and were not downloaded at all. Smear photometry has been used to recover the brightest otherwise-unobserved stars in nominal \kepler (Pope et al., in prep.), and this can also be done in K2, although in the latter case the sample is much smaller due to competition with halo apertures and the systematics correction is more difficult. A natural extension of both pieces of work would be to produce smear light curves of all bright stars without halo apertures in K2, which would finally make the \kepler extended mission magnitude-complete at the bright end.

>>>>

%% Mark off the abstract in the ``abstract'' environment. 
\begin{abstract}
While the \kepler\ Mission was designed to look at tens of thousands of faint stars ($V \gtrsim 12$), brighter stars, which saturate the detector, are important because they can be and have been observed by other instruments at very high signal-to-noise ratio. By considering the unsaturated scattered light `halo' around these stars we can and do retrieve precise light curves of most of the brightest stars in \ktwo fields from Campaign~4 onwards. The halo method is highly agnostic about the cause and form of systematics and we show that it nevertheless it is effective at extracting light curves from both normal and saturated stars. The key methodology is to optimize the weights of a linear combination of pixel time series with respect to an objective function. We test a range of such objective functions, finding that generalizations of Total Variation perform well on both saturated and unsaturated \ktwo targets. Applying this to the bright stars across the \ktwo Campaigns, this reveals stellar variability ubiquitously, including effects of stellar pulsation, rotation, and binarity. Here we describe our pipeline, and present a catalogue of the bright stars studied, with classifications and parametrizations of their variability and remarks on interesting objects. These light curves are publicly available as a High Level Science Product from the Mikulski Archive for Space Telescopes (MAST). \href{https://github.com/benjaminpope/k2halo}{\color{linkcolor}\faGithub} % git link shamelessly borrowed from Luger!
\end{abstract}

@@

% describe saturation
The \kepler detector saturates at a magnitude of $K_p \sim 11.3$ in both long- (30\,min) and short (1\,min)-cadence data, as these both represent sums of 6~s exposures \citep{Gilliland2010}. For objects brighter than this, additional photons create excess electrons which `bleed' into adjacent pixels in both directions along the column containing the bright star. Simple aperture photometry (SAP) -- adding all the flux contained in a window around the bleed column -- has recovered light curves with precisions close to the photon noise limit of stars as bright as 16~Cyg~AB, $\theta$~Cyg and RR~Lyr \citep[e.g.][]{Kolenberg2011,2013MNRAS.433.1262W,Guzik2016}. In the nominal \kepler mission this was not attempted for all such bright stars, and in \ktwo, the several-pixel spacecraft motion significantly increased the size of the required apertures for SAP photometry of very saturated stars, while also making their systematics more difficult to deal with. While the second-version pixel-level-decorrelation (PLD) pipeline EVEREST~2.0 was able to correct systematics in saturated SAP photometry \citep{everest2}, this is rendered impossible for the very brightest stars whose bleed columns may run to the edge of the detector. Furthermore, bandwidth constraints meant that pixel data were not downloaded for many bright targets in \ktwo.

% briefly intro the halo method and halo literature
In order to recover precise photometry of the brightest stars in \ktwo, we have therefore developed two main approaches, `smear' and `halo' photometry. Smear photometry \citep{Pope2016} uses collateral `smear' calibration data to obtain a 1-D spatial profile with $\sim 1/1000$ of the flux on each CCD. This can be processed recover light curves of stars which were not necessarily conventionally targeted and downloaded with active pixels, because smear data are recorded for all columns. The main disadvantage of this method is that it confuses all stars in the same column, which means that in crowded fields smear light curves tend to be significantly contaminated.

% halo
The more precise method of halo photometry, which is the subject of this paper, uses the broad `halo' of scattered light around a saturated star to recover relative photometry, by constructing a light curve as a linear combination of individual pixel time series and minimizing a Total Variation objective function (TV-min). It has been employed for example on the Pleiades \citep{White2017} and the brightest-ever star on \kepler silicon, Aldebaran \citep{Farr2018}, recovering photometry with a precision nearly that of normally-obtained \ktwo observations of unsaturated stars. Unlike smear, this requires downloading data out to a 12--20~pixel radius around each star, and has accordingly only been possible for stars that were specifically targeted with apertures optimized for this method and for a small number of other stars for which this is fortuitously the case. The pixel requirements for this are sufficiently low that, with the help of the \ktwo Guest Observer office, such apertures were obtained for most of the bright targets from Campaign~4 onwards.

% describe contents of paper: code, catalog, categorization, conclusions
In this Paper we describe numerical experiments testing the TV-min method and extending it to generalizations with different exponents and timescales. We show that the method as previously employed applying standard TV-min is suboptimal, and gain a modest improvement from taking finite differences close to the timescale of \ktwo thruster firings. 
We go on to present complete catalog of long-cadence \ktwo halo light curves, which we have made publicly available. We have employed halo photometry on all stars targeted with appropriate apertures, and have done a preliminary characterization of interesting astrophysical variability. We also document the main changes in the halo data reduction pipeline, \texttt{halophot}, with respect to previous releases. These include oscillating red giants, pulsating and quiet main sequence stars, and eclipsing binaries, many of which are among the brightest objects of their type to have been observed with space photometry. We hope that this diverse catalog of light curves will be useful for an equally-diverse range of astrophysical investigations.

@@

@@

\begin{equation}
    Q_{k,l} \equiv {\sum_i{|f_i - f_{i-\delta}|^k}},
\end{equation}

@@

\begin{align}
\forall_j w_j &> 0\\
\sum_{i=1}^{N} f_i  &= N.
\end{align}

This is a classic convex optimization program with constraints, which we solve with the \texttt{scipy} \citep{scipy} L-BFGS-B nonlinear optimization code \citep{lbfgsb}. $Q_{k,l}$ has analytic derivatives with respect to $w_j$ \citep[calculated with \texttt{autograd};][]{autograd}, and it is therefore extremely fast to optimize and converges well on a global solution. In practice, for computational reasons we optimize over parameters $\tilde{w}_j$ such that $w_j = \text{softmax}(\tilde{w}_j$, as this satisfies the constraint that $\forall_j w_j > 0$, and while this also constrains their sum to be unity, we renormalize $f$ to satisfy its normalization constraint before calculating the objective function and this additional constraint is removed again.

The objective function $Q_{k,l}$ is the $L_k$ norm on a `lagged' finite difference with a lag parameter $\delta$. For $k = 1$ and $\delta = 1$, $Q_{1,1}$ is the standard Total Variation objective (TV) used in previous halo papers, and can be seen as the L1 norm on the derivative of $f$ or as a discrete approximation to its arc length. The L2 Variation (L2V) with $k=2$ is sometimes referred to in image processing literature as the `smoothness' regularizer, as it seeks to penalize large gradients without necessarily making them sparse. The lag parameter $\delta$ allows for flexibility in modelling systematics occurring at different timescales from cadence-to-cadence, and we investigate its effects below. The order parameter $k$ allows for flexibility in how sensitive we are to normally-distributed versus long-tailed noise. For convenience in the rest of this paper, we will refer to the $k=1$ case as TV, the $k=2$ case as L2V, and the $k=3$ case as L3V.

@@

In order to choose the values for $k$ and $\delta$ in our objective function, we have selected the quiet star 36~Ophiuchi, the lowest-mass main sequence star in the sample of stars with halo apertures, and one with very little high frequency variability detected or predicted, which was also observed at short cadence. We choose the 6.5~hour Combined Differential Photometric Precision \citep[CDPP,][]{cdpp} as implemented in \texttt{lightkurve} \citep{lightkurve} as a proxy for the `noise' in a lightcurve, with lower being better. We calculate halo lightcurves and their CDPPs for $k \in \{1,2,3\}$, and $\delta \in [1,50]$ for long cadence and for various valies of $\delta \in [1,2500]$ for short cadence data. The results are displayed in Figure~\ref{cdpps}. We find that for long cadence data, $k=1$ (TV) and a lag $\delta=10$ provide the best CDPP. This is unsurprising: that this is one cadence shorter than the 12~cadence thruster firing period. In this context we can understand the optimum as suppressing systematics on the same timescale as they occur. On the other hand, for short cadence data, performance at short lags is very poor but the method performs similarly for $k \in \{1,2\}$ with slow improvement with larger $\delta$, and performs very poorly for $k=3$ at all lags.

% \begin{figure*}
% \plottwo{lag_lc_cdpp.pdf}{lag_sc.pdf}
% \caption{Behaviour of long cadence (left) and short cadence (right) 6.5~hour CDPP for the quiet dwarf 36~Ophiuchi as a function of lag parameter $\delta$. For long cadence this shows a minimum for L1 norm and $\delta=10$, i.e. for objective function $Q_{1,10}$. For short cadence CDPP continuously improves for higher lags and shows no strong differences between L1 and L2~norms, while L3 performs poorly.}
% \label{cdpps}
% \end{figure*}

\begin{figure*}
\plottwo{lag_lc_cdpp.pdf}{lag_lc_sys.pdf}
\caption{Behaviour of long cadence 6.5~hour CDPP (left) and $4 c/d$ systematics power (right) for the quiet dwarf 36~Ophiuchi as a function of lag parameter $\delta$. CDPP shows a minimum for L1 norm and $\delta=10$, i.e. for objective function $Q_{1,10}$.}
\label{cdpps}
\end{figure*}

\begin{figure}
\plotone{lag_sc.pdf}
\caption{Behaviour of short cadence 6.5~hour CDPP for the quiet dwarf 36~Ophiuchi as a function of lag parameter $\delta$. CDPP continuously improves for higher lags and shows no strong differences between L1 and L2~norms, while L3 performs poorly.}
\label{cdpps}
\end{figure}

%  improvements
Unlike other methods for calibrating \kepler systematics, other than the value of $\delta$, no knowledge of the spacecraft motion or the behaviour of an ensemble of other stars is used to inform our algorithm. The method is a self-calibration that is independent of the details of the systematics it is calibrating, operating on the assumption that a single signal is present across many individual time series which otherwise are contaminated by noise. It is therefore likely that significant improvements can be made to the method by including cotrending basis vectors with mean zero and whose weights are allowed to be negative, which would represent systematics which are common to all pixels in the halo aperture and therefore masquerade as signal. Any linear combination of convex objective functions is itself convex, and future extensions to the method could enforce combinations of different lags and orders to better represent systematics occurring on different timescales (e.g. thruster firings, red noise) and with different levels of smoothness.

@@

% do k2 smear after that
The sample of K2 bright stars presented here only includes those with halo apertures, but while some others are available conventionally, many were not assigned target pixels and were not downloaded at all. Smear photometry has been used to recover the brightest otherwise-unobserved stars in nominal \kepler (Pope et al., in prep.), and this can also be done in K2, although in the latter case the sample is much smaller due to competition with halo apertures and the systematics correction is more challenging. A natural extension of both pieces of work would be to produce smear light curves of all bright stars without halo apertures in K2, which would finally make the \kepler extended mission magnitude-complete at the bright end.
